---
title: "Set over Set Simulation"
author: "Jason Laso"
date: "October 11, 2017"
output: 
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    theme: cerulean
---


##### In this simulation, I will attempt to answer the following question: given that a player is dealt a pocket pair, what are the odds that he is oversetted by another player?

## User Functions

##### This function converts list of randomly sampled opponent cards into a matrix of hands. When a sample of opponent cards are taken during the siumulation, it returns them all in one vector. This function will then take that vector, and split the vector of those cards into a matrix of pairs (representing each opponent). This will make it a lot easier to analyze if any of the opponents flopped a set or not.

```{r}
convertHands = function(opp.cards, players){
  #takes a vector of opponent cards and the number of players at the table (incl hero)
  
  opp.hands=matrix(nrow=players-1, ncol=2)
  for(x in 0:(players-2)){
    opp.hands[x+1, ] = (opp.cards[(1:2)+(2*x)])
  }
  return(opp.hands)
}
```

##### This function will take the matrix of opponent hands generated, and then checks to see if any are pocket pairs. It returns a logical vector of true/falses stating if each set of hole cards is or is not a pair.
```{r}
is.pocket.pair = function(opp.hands){
  #takes a matrix of opponent hands (use ConvertHands function)
  
  is.pair = c()
  
  #Takes all combos of opp hands and checks to see if both belong to any of the individual ranks of cards in the deck.
  for(i in 1:nrow(opp.hands)){
    for(rk in 1:13){
      
      if(all(opp.hands[i,] %in% rank[rk,])){
        is.pair[i] = T
        break
      }
      is.pair[i] = F 
    }
    
  }
  
  return(is.pair)
}
```

##### This function takes a matrix of all the other pocket pairs of the opponents at the table, and then return their index name in the deck we will generate (i.e. a hand 1-2 will return the index for all deuces in the deck of cards). Ths index will allow us to more easily check if the hole cards flopped a set in the simulation.
```{r}
opp.pair.rank = function(opp.pairs){
  #takes a matrix of all pocket pairs dealt to opponents
  
  opp.pair.rk = matrix(ncol=1, nrow = nrow(opp.pairs))
  #checks all pairs opponents may have
  for(pair in 1:nrow(opp.pairs)){
    
    #for all ranks of cards in deck
    for(i in 1:13){
      
      if(all(opp.pairs[pair,] %in% rank[i,])){
        opp.pair.rk[pair,] = rownames(rank)[i]
        break
      }
    }
    
  }
  return(opp.pair.rk)
}
```

##### This is the most important function of all because it determines the result of each iteration of the simulation. This function determines what the set over set situation is (no sets at all, hero is only set, hero was the overset, hero was the underset). 
```{r}
set.check = function(flop, opp.pair.rk, hero.pair.rk){
  #takes the flop, the index ranks of the opponent pocket pairs, and the index rank of the hero's pocket pair
  
  if(any(flop %in%  rank[hero.pair.rk,])){ #if hero flopped set 
    
    for(x in opp.pair.rk){   
      
      if(any(flop %in%  rank[x,])){#opp did flop a set
        
        if(max(rank[x,]) < max(rank[hero.pair.rk,])){ #if opponent's set < hero's set
          result = "was.the.overset"
          return(result)
          break
          
        } else{ #opponent's set > hero's set
          result="was.the.underset"
          return(result)
          break
          
        }
      } #opp didn't flop  a set
    } #all opponents didn't flop a set
    
    result = "only.set"
    return(result)
    
  }
  result = "no.set"
  return(result)
}
```

##Simulation Setup

##### This step just creates a deck where the numbers represent ranks of cards (1:4 are deuces, 5:8 are threes, etc.). It also names an index in the form "rk.2", "rk.3", etc.
```{r}
rank=matrix(nrow = 13, ncol=4)

for(x in 0:12){
  rank[x+1,] = ((1:4)+(4*x))
}

rownames(rank) = c(paste("rk", seq(2,14,1), sep="."))
```

##### Set the number of players at the table (including hero). 9-handed is the default.
```{r}
players = 9
```

##### Set the simulation parameters, including the seed (to guarantee reproducability) and the number of iterations. Since set over sets are likely infrequent, I'm setting the iterations to a high number to try and get more positive trials.
```{r}
set.seed(1)
iter=500000
```

##Simulation

##### This is the simulation, which randomly picks out a pocket pair for the hero, deals two random cards to each opponent at the table, checks if any of the opponents have a pocket pair, and then if there is at least one other pair it will evaluate if a set over set situation occurred.
  
```{r, eval=FALSE, include=FALSE}
results = c()


for(i in 1:iter){
  
  card = sample(rownames(rank),1)  #hole card rank
  pocket.pair = sample(rank[card,],2) #create a pair of your hole card
  deck = subset(1:52, !(1:52 %in% pocket.pair)) #remove your pair from the deck
  
  opp.cards = sample(deck, (players-1) * 2) #get all the opponents hole cards as a vector
  opp.hands = convertHands(opp.cards, players) #convert the vector of cards into matrix of player hands
  
  deck = subset(deck, !(deck %in% opp.hands)) #remaining cards in deck after all opponents hole cards taken out
  flop = sample(deck, 3, replace=F) #take a random flop
  
  opp.pair.check = is.pocket.pair(opp.hands) #checks if any other opponents have a pocket pair
  
  #if nobody else has a pair, then there's no need to check if there is set over set in the next step
  if(any(opp.pair.check) == F){
    
    if(any(flop %in%  rank[card,])){ #if the hero flopped a set by himself
        results[i] = "only.set"
        next
    } else{ #if nobody floppped a set
        results[i] = "no.set"
        next 
    }
  }
  
  opp.pairs = matrix(opp.hands[opp.pair.check], ncol=2) #get the matrix of only the opponent pocket pairs
  opp.pair.rk = opp.pair.rank(opp.pairs) #get the names of the ranks of the opponent pocket pairs for indexing
  
  results[i] = set.check(flop, opp.pair.rk, card) #if at least one opponent has a pair, then this step uses our function to check if a set over set happened
  
  print(i) #prints the iteration number after completion
  
} #ends simulation

```

##Results
  
##### Table of results

```{r, eval=FALSE, include=FALSE}
party = data.frame(table(results), prop.table(table(results)) ) %>% select(1,2,4)
colnames(party)[3] = "Freq.pct"

party
```

>       results   Freq Freq.pct
>       no.set 441285 0.882570
>     only.set  56324 0.112648
>     was.overset   1156 0.002312
>     was.underset   1235 0.002470
